Realizing temporal graphs from fastest travel times

### 2023-02-07 ICALP23-RealizingTemporalGraphs.tex
Tuesday 7th February

Graph realization problems are fundamental graph-algorithmic questions where the objective is to (re)construct a graph, 
given some constraints on some of its graph attributes. Among the most well studied such parameters for graph reconstruction 
are the vertex degrees~\cite{cai2000solution,rfgh611961hu,hakimi1962realizability,barNoy2019graphRealizations,barNoy2021selected}, distances~\cite{barNoy2022GraphRealization,barNoy2021composed,hakimi1965distance}, eccentricities~\cite{barNoy2020efficiently,hell2009linear}, connectivity~\cite{fulkerson1960zero,frank1992augmenting,chen1966realization,erdos1960graphs}, etc. 
The motivation for network realization problems stems from both ``verification'' applications where we attempt to reconstruct an existing network of unknown structure based on the outcomes of experimental measurements, and engineering applications related to network design~\cite{augustine2022distributed}.

@@

@@

	When considering the problem of finding shortest paths among vertices, one relies heavily on the optimal substructure of the problem,
	which states that if a shortest path $P$ from vertex $u$ to $v$ traverses the vertex $w$,
	then the shortest path from $u$ to $w$ is a subpath of $P$.
	After studying the problem of fastest temporal paths (\ie temporal paths of smallest duration),
	one can see that this property does not carry over
	(for an example see~\cref{fig:ftpExample}).
	Which makes all the standard intuition and work with these kind of paths, much harder.

@@

In this section we study the hardness of \deltaExact. We first start with defining certain notions, that will be of use when solving the problem.

@@

\begin{proof}
    Let $(P,\lambda)$ be a fastest temporal path from $v_1$ to $v_k$ with the properties from the claim, and let $v_i$ be an arbitrary vertex in $P \setminus \{v_1,v_k\}$.
    Using the properties of fastest paths and the definition of duration, we can rewrite \cref{eq:trav-delays-paths} as follows
    \begin{align*}
        \tau_{v_i}^{v_{i-1},v_{i+1}} & = D_{1,i+1} - D_{1,i} =  d(P_{1,i+1}) - d(P_{1,i}) \\
        & \equiv 
        \left(\lambda(v_{i}v_{i+1}) - \lambda(v_1v_2) + 1\right) -  
        \left(\lambda(v_{i-1}v_{i}) - \lambda(v_1v_2) + 1\right) \pmod \Delta \\
        & \equiv \lambda(v_{i}v_{i+1}) - \lambda(v_{i-1}v_{i}) \pmod \Delta,
        \end{align*}
    which is exactly the definition of $\tau_{v_i}^{v_{i-1},v_{i+1}}$.
\end{proof}

\subsection{ \deltaExact\ is polynomial-time solvable on trees and paths}

Let $D$ be a matrix from \deltaExact, when the underlying graph $G$ of $D$ is a tree on $n$ vertices $\{v_1, v_2, \dots, v_n\}$.
Let $v_i,v_j$ be two arbitrary vertices in $G$. Then we know that there exists a unique (static) path $P$ among them.
Consequently, it follows that the temporal paths $(P_{i,j},\lambda)$ from $v_i$ to $v_j$ and $(P_{j,i},\lambda)$ from $v_j$ to $v_i$ are also unique, up to modulo of the period $\Delta$ of the labeling $\lambda$,
and therefore are the fastest.
Then $D$ is of the following form:
\begin{equation*}
    D_{i,j} =
    \begin{cases}
    0 & \text{if $i = j$}, \\
    1 & \text{if $v_i$ and $v_j$ are neighbours in G}\\
    d(P_{i,j},\lambda) & \text{else}
    \end{cases},
\end{equation*}
where $d(P_{i,j},\lambda)$ is the duration of the (unique) temporal path $(P_{i,j},\lambda)$ from $v_i$ to $v_j$.

@@

Let us observe some properties of the matrix $D$ from \deltaExact, when the underlying graph $G$ of $D$ is a cycle $C_n = \{v_1, v_2, \dots, v_n\}$  on $n$ vertices.
By the definition, each vertex is on distance $0$ from itself. This corresponds with all diagonal entries of $D$ being $0$.
Now, let us observe that each vertex $v_i$ has exactly two neighbours in $C_n$, namely $v_{i-1}$ and $v_{i+1}$,
therefore for all $i\in [n]$ we set $D_{i,i-1} = D_{i, i+1} = 1$, where indices are taken modulo $n$.
This results in the upper and lower diagonal of $D$ having all $1$'s, together with $D_{1,n} = D_{n,1} = 1$.
The matrix $D$ is of the following form
\begin{equation} \label{eq:D-matrixForCycles}
  D = 
  \begin{bmatrix}
    0 & 1 &  &  &  & 1\\
    1 & 0 & 1 &  &  & \\
    & 1 & 0 & 1 & & \\
    & & \ddots & \ddots & \ddots &   \\
    &   & &1 &  0 & 1\\ 
    1 &  &  &  &1 &  0 
  \end{bmatrix},
\end{equation}
where the empty entries consists of positive integers different than $1$.

@@

    Suppose for the contradiction that the statement of the claim does not hold. Then there exists a vertex $v_i \in C_n$, for which we did not calculate its travel delay.
    Let $v_j \in C_n \setminus \{v_i\}$ be an arbitrary vertex. 
    Note that the only time we cannot calculate the travel delay at vertex $v_i$, when considering vertex $v_j$, is in this case when $v_i$ is one of the maximum elements in $D_j$, \ie $v_i$ is a vertex that is on a maximum duration from $v_j$.
    This has to hold for all vertices $v_j$, therefore $v_i$ has to be on the maximum distance from every vertex $v_j \in C_n \setminus \{v_i\}$.
    It then also has to hold for vertices $v_{i-1}, v_{i+1}$, that are neighbours of $v_i$.
    We know that in this case $d^+(i-1,i) = 1, d^-(i+1,i) = 1$.
    Since $v_i$ has to be on the maximum duration from both of them 
    (\ie is one of the two maximum values $m^1_{i-1}, m^2_{i-1}$ for vertex $v_{i-1}$ and one of the two maximum values $m^1_{i+1}, m^2_{i+1}$ for vertex $v_{i+1}$),
    we know that 
    \begin{align}\label{eq:cycle-travel-delays-all}
        & d^+(i+1,i-1) < d^-(i+1,i-1) & \text{and}\\
        & d^-(i-1,i+1) < d^+(i-1,i+1).
    \end{align}
    If this would not hold, the fastest path would go through $v_i$ and we would be able to calculate the travel delay at $v_i$.

@@

\section{\deltaExact\ is NP-hard}

\begin{theorem}
	\deltaExactLong\ is \NP-hard.
\end{theorem}

@@

	\draw [dotted] (x3f) [out=30,in=150] to (x4t);
	\draw [dotted] (x3f) [out=30,in=150] to (x4f);
\end{tikzpicture}
}
\caption{Example of the temporal graph $(G_\phi,\lambda)$ from the NP-hardness reduction, 
	where the 3-SAT formula $\phi$ is of form $\phi = (x_1 \vee \overline{x_2} \vee x_3) \wedge (\overline{x_1} \vee x_2 \vee x_4)$.
For the simplicity we draw edges between vertices $x_i^T$ and $x_j^F$ (where $i \neq j$) with a dotted line.
Presented is the labeling of $G_\phi$, corresponding to the assignment $x_1=x_2=\textsc{True}$ and $x_3,x_4=\textsc{False}$,
where all unlabeled edges get the label $2$.
\label{fig:NP-example}}
\end{figure}
%
Now we specify the distances in the matrix $D^\phi$ between all vertex pairs.
For the sake of simplicity we write $D^\phi _{u,v}$ as $d(u,v)$,
where $u,v$ are two vertices of $G$. 
We start by setting the value of $d(u,v) = 1$ where $u$ and $v$ are two adjacent vertices in $G$.
\begin{itemize}
    %\item For each variable $x$ in $\phi$ we specify the following distances between the non-adjacent variable vertices:
    %$d(x_1,x_2)=2$. 
    %\item For pair of variable $x,y$ in $\phi$ with $x\neq y$ we specify the following 
    %$d(x_1,y^T)=d(x_1,y^F)=3$, $d(x_2,y^T)=d(x_2,y^F)=2$.
    \item For each variable $x_i$ in $\phi$ and the super vertex $v$
    we specify the following distances:
    $d(x_i,v)=2$ and $d(v,x_i)= \Delta $. %, $d(x_2,v)=3$.
    \item For each clause $c$ in $\phi$ and the super vertex $v$
    we specify the following distances:
    $d(c,v)=2$ and $d(v,c)= \Delta - 1$.
    \item Let $x_i$ be a variable that appears in clause $c$, then  we specify the following distances:
    $d(c,x_i)=2$ and $d(x_i,c)=\Delta$.
    If $x_i$ appears non-negated in $c$ we specify the following distances:
    $d(c,x_i^F)=2$ and $d(x_i^F,c)=\Delta$.
    If $x_i$ appears negated in $c$ we specify the following distances:
    $d(c,x_i^T)=2$ and $d(x_i^T,c)=\Delta$.
    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we specify the following distances:
    $d(x_i,c)=\Delta$, $d(c,x_i)=2\Delta + 2$
    and
    $d(c,x_i^T)=d(c,x_i^F)=2$, $d(x_i^T,c)=d(x_i^F,c)=\Delta$. 
    \item For each pair of variables $x_i \neq x_j$ in $\phi$ we specify the following distances:
    $d(x_i,x_j)=2\Delta +1$ and
    $d(x_i,x_j^T)=d(x_i,x_j^F)=\Delta + 1$.
    \item For each pair of clauses $c_i \neq c_j$ in $\phi$ we specify the following distances:
    $d(c_i,c_j)= \Delta + 1$.
\end{itemize}
This finishes the construction of an instance $D^\phi$ of \deltaExact\, which can clearly be done in polynomial time. In the remainder we show that $D^\phi$ is a yes-instance of \deltaExact\ if and only if NAE 3-SAT formula $\phi$ is satisfiable.

@@

We construct a labeling $\lambda$ as follows.
\begin{itemize}
    \item All edges incident with a clause vertex $c$ obtain label one.
    %\item For each variable $x$, we set $\lambda(\{x^T,x^F\})=1$.
    \item If variable $x_i$ is set to \textsc{True}, we set $\lambda(x_i^F v)=3$.
    \item If variable $x_i$ is set to \textsc{False}, we set $\lambda(x_i^T v)=3$.
    \item We set the labels of all other edges to two.
\end{itemize}
For an example of the constructed temporal graph see~\cref{fig:NP-example}.
We now verify that all distances are realized.
\begin{itemize}
    \item For each variable $x_i$ in $\phi$ we have to check that $d(x_i,v)=2$
    and $d(v,x_i)=\Delta$.

@@

    Let $x_k$ be a variable that appears non-negated in $c_i$
    and $x_\ell$ the variable that appears non-negated in $c_j$
    (all other cases are symmetric).
    There is a path of length three from $c_i$ to $c_j$
    that passes first through vertex $x_k^T$ and then through vertex $x_\ell^T$.
    Therefore the temporal path from $c_i$ to $c_j$
    uses the edges $x_k^T c_i, x_\ell ^T c_j, x_k^T x_\ell ^T$, with labels $1,2,1$, respectively,
    which proves the desired equality.
    Note also that this is the shortest path between $c_i$ and $c_j$, and that the first and the last edge must have the label $1$,
    therefore it follows that this is the fastest temporal path.
\end{itemize}
%
\end{proof}

\section{W[1]-hardness}
\begin{theorem}\label{thm:W1wrtFVS}
    \deltaExactLong\ is W[1]-hard when parameterized by the feedback vertex number.
\end{theorem}
\begin{proof}
    We present a parameterized reduction from the W[1]-hard problem \textsc{Multicolored Clique} parameterized by the number of colors~\cite{fellows2009multipleinterval}.  Here, given a $k$-partite graph $H=(W_1\uplus W_2 \uplus\ldots\uplus W_k, F)$, we are asked whether $H$ contains a clique of size $k$. If $w\in W_i$, then we say that $w$ has \emph{color} $i$. W.l.o.g.\ we assume that $|W_1|=|W_2|=\ldots=|W_k|=n$ and that every vertex has at least one neighbor of every color. 
    Furthermore, for all $i\in[k]$, we assume the vertices in $W_i$ are ordered in some arbitrary but fixed way, that is, $W_i=\{w^i_1,w^i_2,\ldots,w^i_n\}$.
    Let $F_{i,j}$ with $i<j$ denote the set of all edges between vertices from $W_i$ and $W_j$. We assume w.l.o.g.\ that $|F_{i,j}|=m$ for all $i< j$.
    Furthermore, for all $i<j$ we assume that the edges in $F_{i,j}$ are ordered in some arbitrary but fixed way, that is, $F_{i,j}=\{e^{i,j}_1,e^{i,j}_2,\ldots,e^{i,j}_m\}$.

We give a reduction to a variant of \deltaExactLong\ where the period $\Delta$ is infinite (that is, there are no periods) and we allow $D$ to have infinity entries, meaning that the two respective vertices are not temporally connected. At the end of the proof we show how to obtain the result for finite $\Delta$ and a distance matrix $D$ that only has integer entries.

    Given an instance $H$ of \textsc{Multicolored Clique}, we construct an instance $D$ of \deltaExactLong\ (with infinity entries and no periods) as follows. To ease the presentation, we first describe the underlying graph $G$ that is implicitly defined by the entries $D_{v,v'}=1$, that is, the pairs of vertices $v,v'$ that should be connected by a temporal path of duration one, meaning that there needs to an edge connecting the two vertices. Afterwards, we describe the remaining entries of $D$.
    We will construct $G$ using several gadgets.

    \subparagraph{Edge Selection Gadget.} We first introduce an \emph{edge selection gadget for color combination $i,j$} with $i<j$. We start with describing the vertex set of the gadget.
    \begin{itemize}
        \item A set $X_{i,j}$ of vertices $x_1, x_2, \ldots, x_m$.
        %\item A set $Y_{i,j}$ of vertices $y_1, y_2, \ldots, y_m$.
        %\item A set $Z_{i,j}$ of vertices $z_1, z_2, \ldots, z_m$.
        \item Vertex sets $U_1, U_2, \ldots, U_m$ with $4n+1$ vertices each, that is, $U_\ell=\{u^\ell_0, u^\ell_1, u^\ell_2,\ldots, 
        %u^\ell_n, u^\ell_{n+1},\ldots, u^\ell_{4n-1}, 
        u^\ell_{4n}\}$ for all $\ell\in[m]$. 
        %Let $U_{i,j}=\bigcup_{\ell\in[n]} U_\ell$.
        \item Two special vertices $v_{i,j}^\star,v_{i,j}^{\star\star}$.
    \end{itemize}
    The gadget has the following edges.
    \begin{itemize}
        \item For all $\ell\in [m]$ we have edge $\{x_\ell,v_{i,j}^\star\}$, $\{v_{i,j}^\star,u^\ell_0\}$, and $\{u^\ell_{4n},v_{i,j}^{\star\star}\}$.

@@

    \subparagraph{Connector Gadget.} Next, we describe \emph{connector gadgets}. Intuitively, these gadgets will be used to connect many vertex pairs by short paths, which will make arguing about possible labelings in yes-instances much easier. Connector gadgets consist of six vertices $\hat{v}_0,\hat{v}_0',\hat{v}_1,\hat{v}_2,\hat{v}_3,\hat{v}_3'$. 
    %Each connector gadget is associated with four sets $A,A',B,B'$ containing vertices of other gadgets. 
    Each connector gadget is associated with two sets $A,B$ with $B\subseteq A$ containing vertices of other gadgets. 
    Let $V^\star$ denote the set of all vertices from all edge selection gadgets and all validation gadgets.
    The sets $A$ and $B$ will only play a role when defining the distance matrix $D$ later. Informally speaking, vertices in $A$ should reach all vertices in $V^\star$ quickly through the gadget, except the ones in $B$.
    We have the following edges. 
    \begin{itemize}
        \item Edges $\{\hat{v}_0,\hat{v}_1\},\{\hat{v}_0',\hat{v}_1\},\{\hat{v}_1,\hat{v}_2\},\{\hat{v}_2,\hat{v}_3\},\{\hat{v}_2,\hat{v}_3'\}$.
        \item An edge between $\hat{v}_1$ and each vertex in $V^\star$.
        \item An edge between $\hat{v}_2$ and each vertex in $V^\star$.
    \end{itemize}
    We add two connector gadgets for each edge selection gadget and two connector gadgets for each validation gadget.

@@

\subparagraph{Distance Matrix.} We proceed with describing the distance matrix $D$. For a more convenient presentation, we use the notation $d(v,v'):= D_{v,v'}$. For all vertices $v,v'$ that are neighbors in $G$ we have that $d(v,v')=1$ and $d(v',v)=1$.

>>>>

Graph realization problems are fundamental graph-algorithmic questions where the objective is to (re)construct a graph, 
given some constraints on some of its graph attributes. Among the most well studied such parameters for graph reconstruction 
are the vertex degrees~\cite{cai2000solution,rfgh611961hu,hakimi1962realizability,barNoy2019graphRealizations,barNoy2021selected,erdos1960graphs}, distances~\cite{barNoy2022GraphRealization,barNoy2021composed,hakimi1965distance}, eccentricities~\cite{barNoy2020efficiently,hell2009linear,behzad1976eccentric,lesniak1975eccentric}, connectivity~\cite{fulkerson1960zero,frank1992augmenting,chen1966realization,frank1994connectivity,frank1970connectivity,gomory1961multi}, etc. 
The motivation for network realization problems stems from both ``verification'' applications where we attempt to reconstruct an existing network of unknown structure based on the outcomes of experimental measurements, and engineering applications related to network design~\cite{augustine2022distributed,grotschel1995design}.

@@

Sections for ICALP:
\begin{enumerate}
    \item Intro 2 PAGES (GEORGE + PAUL)
    \item Prelims 1 PAGE
    \item Hardness (HENDRIK SHORTENS)
    \begin{enumerate}
        \item NP 1/2 PAGE (APPENDIX)
        \item W1 (3 PAGES)
    \end{enumerate}
    \item Algos (NINA SHORTENS)
    \begin{enumerate}
        \item Tree ALGO 2 PAGE
        \item FPT ALGO (6 PAGES)
        \item Cylce ALGO (APPENDIX)
    \end{enumerate}
    \item Conclusion (?) 1/2 PAGE
\end{enumerate}

	When considering the problem of finding shortest paths among vertices, one relies heavily on the optimal substructure of the problem,
	which states that if a shortest path $P$ from vertex $u$ to $v$ traverses the vertex $w$,
	then the shortest path from $u$ to $w$ is a subpath of $P$.
	After studying the problem of fastest temporal paths (\ie temporal paths of smallest duration),
	one can see that this property does not carry over
	(for an example see~\cref{fig:ftpExample}).
	Which makes all the standard intuition and work with these kind of paths, much harder.
    This also explains why our FPT algorithm, w.r.t the feedback edge number, is so complex. Usually, the feedback edge number parameter (denote it $k$) provides an easy algorithm, 
    where one first determines $O(k)$ special vertices,
    guesses desired structures among them (\ie iterates through all possibilities) and then simply extends the solution to the rest of the graph.
    An example with this approach, on temporal graphs, can be observed in \cite{casteigts2021finding} where the authors provide an FPT algorithm that determines all temporal paths, under certain constraints.
    In our case the similar approach cannot work, as we cannot at any point calculate some partial labeling $\lambda$ of a delta periodic graph $(G,\lambda)$,
    that could at the end be extended to the whole $G$.

@@

@@

In this section 
we show that in the case when the underlying graph $G$ of an instance $(D, \Delta)$ of \deltaExact\ is a tree or a cycle,
we can determine desired $\Delta$-periodic labeling $\lambda$ of $G$ in polynomial time.
We first start with defining certain notions, that will be of use when solving the problem.

@@

\begin{proof}
    Let $(P,\lambda)$ be a fastest temporal path from $v_1$ to $v_k$ with the properties from the claim, and let $v_i$ be an arbitrary vertex in $P \setminus \{v_1,v_k\}$.
    Using the properties of fastest paths and the definition of duration, we can rewrite \cref{eq:trav-delays-paths} as follows
    \begin{align*}
        \tau_{v_i}^{v_{i-1},v_{i+1}} & = D_{1,i+1} - D_{1,i} =  d(P_{1,i+1}) - d(P_{1,i}) \\
        & \equiv 
        \left(\lambda(v_{i}v_{i+1}) - \lambda(v_1v_2) + 1\right) -  
        \left(\lambda(v_{i-1}v_{i}) - \lambda(v_1v_2) + 1\right) \pmod \Delta \\
        & \equiv \lambda(v_{i}v_{i+1}) - \lambda(v_{i-1}v_{i}) \pmod \Delta,
        \end{align*}
    which is exactly the definition of $\tau_{v_i}^{v_{i-1},v_{i+1}}$.
\end{proof}
All of the above is used in the design of the following polynomial-time algorithms.
\subsection{ \deltaExact\ is polynomial-time solvable on trees and paths}
%
Let $D$ be a matrix from \deltaExact, when the underlying graph $G$ of $D$ is a tree on $n$ vertices $\{v_1, v_2, \dots, v_n\}$.
Let $v_i,v_j$ be two arbitrary vertices in $G$. Then we know that there exists a unique (static) path $P$ among them.
Consequently, it follows that the temporal paths $(P_{i,j},\lambda)$ from $v_i$ to $v_j$ and $(P_{j,i},\lambda)$ from $v_j$ to $v_i$ are also unique, up to modulo of the period $\Delta$ of the labeling $\lambda$,
and therefore are the fastest.
Then $D$ is of the following form:
\begin{equation*}
    D_{i,j} =
    \begin{cases}
    0 & \text{if $i = j$}, \\
    1 & \text{if $v_i$ and $v_j$ are neighbours in G}\\
    d(P_{i,j},\lambda) & \text{else}
    \end{cases},
\end{equation*}
where $d(P_{i,j},\lambda)$ is the duration of the (unique) temporal path $(P_{i,j},\lambda)$ from $v_i$ to $v_j$.

@@

Let us observe some properties of the matrix $D$ from \deltaExact, when the underlying graph $G$ of $D$ is a cycle $C_n = \{v_1, v_2, \dots, v_n\}$  on $n$ vertices.
By the definition, each vertex is on distance, and therefore also duration $0$ from itself. This corresponds with all diagonal entries of $D$ being $0$.
Now, let us observe that each vertex $v_i$ has exactly two neighbours in $C_n$, namely $v_{i-1}$ and $v_{i+1}$,
therefore for all $i\in [n]$ we set $D_{i,i-1} = D_{i, i+1} = 1$, where indices are taken modulo $n$.
This results in the upper and lower diagonal of $D$ having all $1$'s, together with $D_{1,n} = D_{n,1} = 1$.
The matrix $D$ is of the following form
\begin{equation} \label{eq:D-matrixForCycles}
  D = 
  \begin{bmatrix}
    0 & 1 &  &  &  & 1\\
    1 & 0 & 1 &  &  & \\
    & 1 & 0 & 1 & & \\
    & & \ddots & \ddots & \ddots &   \\
    &   & &1 &  0 & 1\\ 
    1 &  &  &  &1 &  0 
  \end{bmatrix},
\end{equation}
where the empty entries consists of positive integers different than $1$.

@@

    Suppose for the contradiction that the statement of the claim does not hold. Then there exists a vertex $v_i \in C_n$, for which we did not calculate its travel delay.
    Let $v_j \in C_n \setminus \{v_i\}$ be an arbitrary vertex. 
    Note that the only time we cannot calculate the travel delay at vertex $v_i$, when considering vertex $v_j$, is in this case when $v_i$ is one of the maximum elements in $D_j$, \ie $v_i$ is a vertex that is on a maximum duration from $v_j$.
    This has to hold for all vertices $v_j$, therefore $v_i$ has to be on the maximum duration from every vertex $v_j \in C_n \setminus \{v_i\}$.
    It then also has to hold for vertices $v_{i-1}, v_{i+1}$, that are neighbours of $v_i$.
    We know that in this case $d^+(i-1,i) = 1, d^-(i+1,i) = 1$.
    Since $v_i$ has to be on the maximum duration from both of them 
    (\ie is one of the two maximum values $m^1_{i-1}, m^2_{i-1}$ for vertex $v_{i-1}$ and one of the two maximum values $m^1_{i+1}, m^2_{i+1}$ for vertex $v_{i+1}$),
    we know that 
    \begin{align}\label{eq:cycle-travel-delays-all}
        & d^+(i+1,i-1) < d^-(i+1,i-1) & \text{and}\\
        & d^-(i-1,i+1) < d^+(i-1,i+1).
    \end{align}
    If this would not hold, the fastest path would go through $v_i$ and we would be able to calculate the travel delay at $v_i$.

@@

\section{Hardness Results}  
\subsection{\deltaExact\ is NP-hard}
In this section we prove that it is NP-hard to determine the $\Delta$-periodic temporal graph $(G,\lambda)$,
for a general instance $(D, \Delta)$ of \deltaExact\,
where $\Delta \geq 3$.

\begin{theorem}
	\deltaExactLong\ is \NP-hard for a general $(D,\Delta)$, where $\Delta \geq 3$.
\end{theorem}

@@

	\draw [dotted] (x3f) [out=30,in=150] to (x4t);
	\draw [dotted] (x3f) [out=30,in=150] to (x4f);
\end{tikzpicture}
}
\caption{Example of the temporal graph $(G_\phi,\lambda)$ from the NP-hardness reduction, 
	where the 3-SAT formula $\phi$ is of form $\phi = (x_1 \vee \overline{x_2} \vee x_3) \wedge (\overline{x_1} \vee x_2 \vee x_4)$.
For the simplicity we draw edges between vertices $x_i^T$ and $x_j^F$ (where $i \neq j$) with a dotted line.
Presented is the labeling of $G_\phi$, corresponding to the assignment $x_1=x_2=\textsc{True}$ and $x_3,x_4=\textsc{False}$,
where all unlabeled edges get the label $2$.
\label{fig:NP-example}}
\end{figure}
%
Now we specify the durations in the matrix $D^\phi$ between all vertex pairs.
For the sake of simplicity we write $D^\phi _{u,v}$ as $d(u,v)$,
where $u,v$ are two vertices of $G$. 
We start by setting the value of $d(u,v) = 1$ where $u$ and $v$ are two adjacent vertices in $G$.
\begin{itemize}
    %\item For each variable $x$ in $\phi$ we specify the following distances between the non-adjacent variable vertices:
    %$d(x_1,x_2)=2$. 
    %\item For pair of variable $x,y$ in $\phi$ with $x\neq y$ we specify the following 
    %$d(x_1,y^T)=d(x_1,y^F)=3$, $d(x_2,y^T)=d(x_2,y^F)=2$.
    \item For each variable $x_i$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(x_i,v)=2$ and $d(v,x_i)= \Delta $. %, $d(x_2,v)=3$.
    \item For each clause $c$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(c,v)=2$ and $d(v,c)= \Delta - 1$.
    \item Let $x_i$ be a variable that appears in clause $c$, then  we specify the following durations:
    $d(c,x_i)=2$ and $d(x_i,c)=\Delta$.
    If $x_i$ appears non-negated in $c$ we specify the following durations:
    $d(c,x_i^F)=2$ and $d(x_i^F,c)=\Delta$.
    If $x_i$ appears negated in $c$ we specify the following duratios:
    $d(c,x_i^T)=2$ and $d(x_i^T,c)=\Delta$.
    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we specify the following duratios:
    $d(x_i,c)=\Delta$, $d(c,x_i)=2\Delta + 2$
    and
    $d(c,x_i^T)=d(c,x_i^F)=2$, $d(x_i^T,c)=d(x_i^F,c)=\Delta$. 
    \item For each pair of variables $x_i \neq x_j$ in $\phi$ we specify the following duratios:
    $d(x_i,x_j)=2\Delta +1$ and
    $d(x_i,x_j^T)=d(x_i,x_j^F)=\Delta + 1$.
    \item For each pair of clauses $c_i \neq c_j$ in $\phi$ we specify the following duratios:
    $d(c_i,c_j)= \Delta + 1$.
\end{itemize}
This finishes the construction of an instance $D^\phi$ of \deltaExact\, which can clearly be done in polynomial time. In the remainder we show that $D^\phi$ is a yes-instance of \deltaExact\ if and only if NAE 3-SAT formula $\phi$ is satisfiable.

@@

We construct a labeling $\lambda$ as follows.
\begin{itemize}
    \item All edges incident with a clause vertex $c$ obtain label one.
    %\item For each variable $x$, we set $\lambda(\{x^T,x^F\})=1$.
    \item If variable $x_i$ is set to \textsc{True}, we set $\lambda(x_i^F v)=3$.
    \item If variable $x_i$ is set to \textsc{False}, we set $\lambda(x_i^T v)=3$.
    \item We set the labels of all other edges to two.
\end{itemize}
For an example of the constructed temporal graph see~\cref{fig:NP-example}.
We now verify that all duratios are realized.
\begin{itemize}
    \item For each variable $x_i$ in $\phi$ we have to check that $d(x_i,v)=2$
    and $d(v,x_i)=\Delta$.

@@

    Let $x_k$ be a variable that appears non-negated in $c_i$
    and $x_\ell$ the variable that appears non-negated in $c_j$
    (all other cases are symmetric).
    There is a path of length three from $c_i$ to $c_j$
    that passes first through vertex $x_k^T$ and then through vertex $x_\ell^T$.
    Therefore the temporal path from $c_i$ to $c_j$
    uses the edges $x_k^T c_i, x_\ell ^T c_j, x_k^T x_\ell ^T$, with labels $1,2,1$, respectively,
    which proves the desired equality.
    Note also that this is the shortest path between $c_i$ and $c_j$, and that the first and the last edge must have the label $1$,
    therefore it follows that this is the fastest temporal path.
\end{itemize}
Lastly, observe that $\Delta$ has to be greater or equal to $3$, as the above constructed labeling $\lambda$ uses values from $\{1,2,3\}$.
\end{proof}

\subsection{W[1]-hardness}
\begin{theorem}\label{thm:W1wrtFVS}
    \deltaExactLong\ is W[1]-hard when parameterized by the feedback vertex number.
\end{theorem}
\begin{proof}
    We present a parameterized reduction from the W[1]-hard problem \textsc{Multicolored Clique} parameterized by the number of colors~\cite{fellows2009multipleinterval}.  Here, given a $k$-partite graph $H=(W_1\uplus W_2 \uplus\ldots\uplus W_k, F)$, we are asked whether $H$ contains a clique of size $k$. If $w\in W_i$, then we say that $w$ has \emph{color} $i$. W.l.o.g.\ we assume that $|W_1|=|W_2|=\ldots=|W_k|=n$ and that every vertex has at least one neighbor of every color. 
    Furthermore, for all $i\in[k]$, we assume the vertices in $W_i$ are ordered in some arbitrary but fixed way, that is, $W_i=\{w^i_1,w^i_2,\ldots,w^i_n\}$.
    Let $F_{i,j}$ with $i<j$ denote the set of all edges between vertices from $W_i$ and $W_j$. We assume w.l.o.g.\ that $|F_{i,j}|=m$ for all $i< j$.
    Furthermore, for all $i<j$ we assume that the edges in $F_{i,j}$ are ordered in some arbitrary but fixed way, that is, $F_{i,j}=\{e^{i,j}_1,e^{i,j}_2,\ldots,e^{i,j}_m\}$.

We give a reduction to a variant of \deltaExactLong\ where the period $\Delta$ is infinite (that is, there are no periods) and we allow $D$ to have infinity entries, meaning that the two respective vertices are not temporally connected. At the end of the proof we show how to obtain the result for a finite $\Delta$ and a matrix $D$ of durations of fastest paths, that only has finite entries.

    Given an instance $H$ of \textsc{Multicolored Clique}, we construct an instance $D$ of \deltaExactLong\ (with infinity entries and no periods) as follows. To ease the presentation, we first describe the underlying graph $G$ that is implicitly defined by the entries $D_{v,v'}=1$, that is, the pairs of vertices $v,v'$ that should be connected by a temporal path of duration one, meaning that there needs to be an edge connecting the two vertices. Afterwards, we describe the remaining entries of $D$.
    We will construct $G$ using several gadgets.

    \subparagraph{Edge Selection Gadget.} We first introduce an \emph{edge selection gadget $G_{i,j}$ for color combination $i,j$} with $i<j$. We start with describing the vertex set of the gadget.
    \begin{itemize}
        \item A set $X_{i,j}$ of vertices $x_1, x_2, \ldots, x_m$.
        %\item A set $Y_{i,j}$ of vertices $y_1, y_2, \ldots, y_m$.
        %\item A set $Z_{i,j}$ of vertices $z_1, z_2, \ldots, z_m$.
        \item Vertex sets $U_1, U_2, \ldots, U_m$ with $4n+1$ vertices each, that is, $U_\ell=\{u^\ell_0, u^\ell_1, u^\ell_2,\ldots, 
        %u^\ell_n, u^\ell_{n+1},\ldots, u^\ell_{4n-1}, 
        u^\ell_{4n}\}$ for all $\ell\in[m]$. 
        %Let $U_{i,j}=\bigcup_{\ell\in[n]} U_\ell$.
        \item Two special vertices $v_{i,j}^\star,v_{i,j}^{\star\star}$.
    \end{itemize}
    The gadget has the following edges.
    \begin{itemize}
        \item For all $\ell\in [m]$ we have edge $\{x_\ell,v_{i,j}^\star\}$, $\{v_{i,j}^\star,u^\ell_0\}$, and $\{u^\ell_{4n},v_{i,j}^{\star\star}\}$.

@@

    \subparagraph{Connector Gadget.} Next, we describe \emph{connector gadgets}. Intuitively, these gadgets will be used to connect many vertex pairs by short paths, which will make arguing about possible labelings in yes-instances much easier. Connector gadgets consist of six vertices $\hat{v}_0,\hat{v}_0',\hat{v}_1,\hat{v}_2,\hat{v}_3,\hat{v}_3'$. 
    %Each connector gadget is associated with four sets $A,A',B,B'$ containing vertices of other gadgets. 
    Each connector gadget is associated with two sets $A,B$ with $B\subseteq A$ containing vertices of other gadgets. 
    Let $V^\star$ denote the set of all vertices from all edge selection gadgets and all validation gadgets.
    The sets $A$ and $B$ will only play a role when defining the matrix $D$ later. Informally speaking, vertices in $A$ should reach all vertices in $V^\star$ quickly through the gadget, except the ones in $B$.
    We have the following edges. 
    \begin{itemize}
        \item Edges $\{\hat{v}_0,\hat{v}_1\},\{\hat{v}_0',\hat{v}_1\},\{\hat{v}_1,\hat{v}_2\},\{\hat{v}_2,\hat{v}_3\},\{\hat{v}_2,\hat{v}_3'\}$.
        \item An edge between $\hat{v}_1$ and each vertex in $V^\star$.
        \item An edge between $\hat{v}_2$ and each vertex in $V^\star$.
    \end{itemize}
    We add two connector gadgets for each edge selection gadget and two connector gadgets for each validation gadget.

@@

\subparagraph{Matrix $D$.} We proceed with describing the matrix $D$, of durations of fastest paths. For a more convenient presentation, we use the notation $d(v,v'):= D_{v,v'}$. For all vertices $v,v'$ that are neighbors in $G$ we have that $d(v,v')=1$ and $d(v',v)=1$.
